HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-14 paper/main.tex
sections 3.1 and 3.2

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of structure-aware\bh{What is ``structure-aware MT''} MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
MEANT \cite{lo2011structured})\bh{Should we just refer to HMEANT, as that is the comparable measure from the MEANT family?}.
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

@@

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
    \node (ROOT) [fill=black, circle] {}
    child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
    child {node (H1) [fill=black,circle] {}
      child {node {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [fill=black,circle] {}
      child {node {} edge from parent[draw=none]}
      child {node {} edge from parent[draw=none]}
      child {node (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node {moved} edge from parent node[right] {$P$}}
      child {node [fill=black,circle] {}
        child {node {to} edge from parent node[right] {$R$}}
        child {node {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation.}
\end{figure}

@@

@@

composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])

>>>>

  Recent interest in semantics-aware MT systems underscores the importance of
  semantic evaluation for Machine Translation (MT) systems. 
  A semantic measure, sensitive to possible semantic discrepencies
  between the translation output and the source, is likely support the construction
  and tuning of semantics-based MT, much in the same way that string-based measures,
  such as BLEU, enabled progress in shallower MT.
  We present a novel human semantic evaluation measure for MT, Human
  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
  Making progress over previous work, the proposed measure takes into account
  a wider range of semantic phenomena and does not rely on semantic annotation
  of the MT output itself, whose interpretation is often unclear.
  We experiment with four language pairs, demonstrating HUME's broad applicability,
  as well as its reliability, reflected in its inter-annotator agreement rates.

@@

These problems were partially addressed by measures that decompose over sub-parts of the evaluated
translation sentence (henceforth, {\it translation}),
most commonly over its words or n-grams,
quantifying the overlap of its sub-parts relative a reference translation.
More recent work on MT evaluation proposed measures
that decompose over semantically defined units,
quantifying the similarity of the output and the reference in terms of
their verb argument structure (the most notable of these measures is
HMEANT \cite{lo2011structured}).
While allowing to localize semantic errors in the translation,
these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
such as inter-clause linkage and nominal argument structures, and thus only provides
a limited perspective on the semantic similarity between the output and reference.
Moreover, assigning semantic structure to MT output, especially when low in quality,
is difficult and arguably ill-defined.

@@

\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
  \node (ROOT) [label={\#7}] [fill=black, circle] {}
  child {node [label={\#1}] {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
  child {node (H1) [label={\#8}] [fill=black,circle] {}
    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    }
    child {node (H2) [label={\#9}] [fill=black,circle] {}
      child {node  {} edge from parent[draw=none]}
      child {node  {} edge from parent[draw=none]}
      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
      child {node [label={\#10}] [fill=black,circle] {}
        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
        edge from parent node[right] {$A$}}
      %child {node {to Paris} edge from parent node[right] {$A$}}
      edge from parent node[left] {$H$}
    };
    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
  \end{tikzpicture}
\caption{\label{fig:ucca_example}
  Sample UCCA annotation. Edge labels mark UCCA categories. \#<index> marks the node
indices.}
\end{figure}

@@

For example, consider the sentence ``After graduation, Tom moved
to NYC'' \figref{fig:ucca_example}. The atomic units are the inidividual words,
nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
  be in English for presentation purposes. In our experiments $t$ is in German,
  Czech, Polish or Romanian.}.
Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
expresses the relation between the graduation and moving Scenes, is marked Bad,
as their relation in $t$ is reversed.

@@

We proceed to detailing how judgements on the semantic units
of the source are aggregated into a composite score.
While future extensive empirical evaluation with HUME, and using it to
tune MT systems may suggest more complex measure,
as a starting point we take a parsimonious approach
and define aggregation of the judgements.

Let $Correct(s,t)$ be the number of atomic units in $s$ marked as Green relative to $t$ plus
the number of structural units in $s$ marked as Adequate relative to $t$.
Let $Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
to $t$. Then HUME's composite score is:

\vspace{-.5cm}
\begin{equation}
  \small
  HUME(s,t) = \frac{Correct(s,t) + 0.5\cdot Partial(s,t)}{|Units(s)|}
\end{equation}

%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
